Automated Topic Naming

### 2011-09-04 ese/ese.tex
more author analysis

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use semi-unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
%160 words
\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}

@@

%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
% this reduces the line length and drops a ref even though we still have a scacchi kickin around
%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
%Their work discusses the source of the requirements and how they are used in the development process. 
% German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
\todo{re-add this related work?}
% XXX vanity or space.. 
%XXXXXXXXXXXX OPTIONAL
% NE 0113 this seems too early to mention 'result'.
% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
% techniques. This supports our result that software qualities are not discussed with the same frequency across projects.

@@

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project,
 the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. 
%It is really important and it is required by both analyses. Plus we can tell them how.}

%\todo[inline]{NEIL: this section should be updated to explain F-measure}
We validate classification performance using the Receiver Operating
Characteristic area-under-curve value~\cite{Fawcett2006861},
abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.

@@

%\todo[inline]{* clarify the mailinglist stuff we did}
%\todo[inline]{This part doesn't make a lot of sense}
To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
%, an English-language ``lexical database'' that contains semantic relations between words,
% including common related forms (similar to word stemming), meronymy and synonymy. 
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

@@

%\todo[inline]{NEIL: this could be a subsubsection}
%\todo[inline]{* the hierarchy of tags was not clear}
%\noindent \textbf{Generating labels automatically} ---

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a match. 
\emph{Unnamed topics} occur where there is no such match. 
This may indicate either a lack of precision in the word-lists, or simply that this topic is not associated with non-functional
requirements.
All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
extracted $20$ topics per period for each project.
% Is this good enough
This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
the project being analyzed, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, we can use the 
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

%\todo[inline,color=green]{this table should have mysql}
\begin{table}
\centering
% % This data came from running  
% %   python check_validity.py
% %  in src/validate
% % and then running
% % tail -n 5 *maxdb*yes_no*
% % tail -n 5 *mysql*yes_no*
% ok the new way is to run lda_relator
% src/validate/lda_relator_counts.sh
% NOTE THESE WERE THE OLD VALUES
% WHY DID THEY CHANGE???
% DUNNO
%Named topics	& 281 & 125 & 328  \\
%Unnamed topics & 139 & 295 & 92   \\

@@

%\todo[inline]{Should we use F-1 as well to satisfy everyone?}

@@

%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? --
% ROC is hot right now and it combines both}
% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +
%it has natural meaningful quality mappings.
% \begin{figure}[t]
% \centering
% \subfloat[ROC]{
%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
% \label{fig:maxdb-unsup-results-roc}
% }
% %\subfloat[F1]{
% %\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
% %\label{fig:maxdb-unsup-results-f1}
% %}
% \caption[]{Performance of unsupervised topic labelling for each NFR per word-list. A) ROC values -- 
% dashed line indicates the performance of a random classifier. B) F1 measure. Possible values range from 0--1.
% %\todo[inline]{better discussion}% and graphical legend would be nice}
% }
% \label{fig:maxdb-unsup-results}
% \end{figure}

@@

%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure*}

@@

\subfloat[PostgreSQL 7.2]{					
\includegraphics[width=\textwidth]{figures/pgsql-timeline}
\label{fig:pgsql-timeline}
}
	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
	Grid cell intensity (saturation) is mapped to label frequency
	relative to the largest label count of \emph{all} NFRs. Grey
	histogram bars indicate label frequency relative to that
	particular NFR's largest label count. Dashed vertical lines
	relate a project milestone (\emph{*Key events*}) to our topic windows. 
% \todo[inline]{better discussion}
}
\label{fig:timelines}
% Timelines created using src/validate/summary_labels.py
\end{figure}

@@

% \label{sec:comparison}
%\XXX{Not sure there's enough here to do that!}
% We observed that MySQL had more topic subsets than MaxDB. 
% MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
% We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
%\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
% \todo[inline]{expand}
%\todo[inline,color=yellow]{expand on this with dates, and reference to visualizations}
%\subsection{Research questions}
%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
%This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
%\todo{add more on PGSQL}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

It should be noted that much of PostgreSQL development originates from
the mailinglist and the developers are often pushing integrations.

@@

@@

\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
of developer comments. 
%The OSS projects investigated were database systems, thus 
%our results might not generalize to other domains. %We only studied database systems.
% \todo[inline]{If we compare with the same domain we don't get to say it generalizes a lot?}

>>>>

\begin{abstract}

Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use semi-unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
%160 words
\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}

@@

%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
% this reduces the line length and drops a ref even though we still have a scacchi kickin around
Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
Their work discusses the source of the requirements and how they are used in the development process. 
 German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
% NE 0113 this seems too early to mention 'result'.
% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
% techniques. This supports our result that software qualities are not discussed with the same frequency across projects.

@@

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project,
 the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?

We validate classification performance using the Receiver Operating
Characteristic area-under-curve value~\cite{Fawcett2006861},
abbreviated \emph{ROC}, and the F-measure, which is the harmonic mean of precision and recall, i.e., $2 * (P * R) / (P + R)$.

@@

To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

@@

%\noindent \textbf{Generating labels automatically} ---

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labelled our topics with an NFR where there was a match between a word in
the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a match. 
\emph{Unnamed topics} occur where there is no such match. 
This may indicate either a lack of precision in the word-lists, or simply that this topic is not associated with non-functional
requirements.
All experiments were run on the datasets for each project (e.g., PostgreSQL, MySQL, MaxDB). LDA
extracted $20$ topics per period for each project.
% Is this good enough
This labelling is \emph{semi-unsupervised} because the corpus is not derived from 
the project being analyzed, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, we can use the 
domain knowledge of software to label relevant topics.

@@

\begin{table}
\centering
% % This data came from running  
% %   python check_validity.py
% %  in src/validate
% % and then running
% % tail -n 5 *maxdb*yes_no*
% % tail -n 5 *mysql*yes_no*
% ok the new way is to run lda_relator
% src/validate/lda_relator_counts.sh
% NOTE THESE WERE THE OLD VALUES
% WHY DID THEY CHANGE???
% DUNNO
%Named topics	& 281 & 125 & 328  \\
%Unnamed topics & 139 & 295 & 92   \\

@@

@@

% \begin{figure}[t]
% \centering
% \subfloat[ROC]{
%  \includegraphics[width=0.45\textwidth]{figures/unsupervised-bar}
% \label{fig:maxdb-unsup-results-roc}
% }
% %\subfloat[F1]{
% %\includegraphics[width=0.45\textwidth]{figures/unsupervised-bar-f1}
% %\label{fig:maxdb-unsup-results-f1}
% %}
% \caption[]{Performance of unsupervised topic labelling for each NFR per word-list. A) ROC values -- 
% dashed line indicates the performance of a random classifier. B) F1 measure. Possible values range from 0--1.

% }
% \label{fig:maxdb-unsup-results}
% \end{figure}

@@

  \label{fig:maxdb-unsup-results}
\end{figure*}

@@

\subfloat[PostgreSQL 7.2]{					
\includegraphics[width=\textwidth]{figures/pgsql-timeline}
\label{fig:pgsql-timeline}
}
	\caption[]{NFR label per period. Each cell represents a 30-day period. %Numbers represent number of topics in that period labelled with that NFR. 
	Grid cell intensity (saturation) is mapped to label frequency
	relative to the largest label count of \emph{all} NFRs. Grey
	histogram bars indicate label frequency relative to that
	particular NFR's largest label count. Dashed vertical lines
	relate a project milestone (\emph{*Key events*}) to our topic windows.

}
\label{fig:timelines}
% Timelines created using src/validate/summary_labels.py
\end{figure}

@@

% \label{sec:comparison}
%\XXX{Not sure there's enough here to do that!}
% We observed that MySQL had more topic subsets than MaxDB. 
% MySQL 3.23 also had more topics and a longer recorded history than MaxDB 7.500. 
% We tagged more MySQL topics with annotations than MaxDB topics, yet both shared similarities. 
% In terms of non-functional requirements both projects had long running trends that focused on \emph{functionality}, \emph{maintainability}, and
%\emph{portability}, yet MaxDB had more of a focus on \emph{efficiency} and \emph{reliability}. 
%\subsection{Research questions}
%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
%This analysis addresses our research questions: %raised in Section \ref{sec:analysis}:

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}

In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

It should be noted that much of PostgreSQL development originates from
the mailinglist and the developers are often pushing integrations. If
we compare the total to each author we find that 1 quarter of the
authors have a similarity of $0.47$ or less. This is depicted in
Figure \ref{fig:authorsim} where we can see that not all authors are
as similar to the global trend as others. Of
course, we found that number of commits had a strong correlation
($0.59$ Pearson) between the similarity to the aggregate and the
number of topics associated with that author.

Our theory is that the less frequent committers perhaps are more
focused and less general, thus their distributions of topics are
different than the main developers who commit code in many different
contexts. Perhaps the lead developers ``wear more hats'' or have more
time to ``wear more hats''.

Given the top 15 PostgreSQL devs we found that proportionally many
devs had a focus on NFR topics. Authors with top proporitions per NFR
were:
\texttt{dennis} for usability; 
\texttt{scrappy} and \texttt{meskes} for portability;
\texttt{inoue} and \texttt{neilc} for efficiency;
\texttt{jurka} and \texttt{joe} for reliability;
\texttt{thomas} and \texttt{weick} for functionality;
and \texttt{ishii} and \texttt{scrappy} for maintainability.
Given these results we found that yes indeed there was often a focus
of $1/3$ to $1/10$ of commits dedicated to a particular NFR.

% Dennis is into usability
% scrappy and meskes portability
% Inoue and neilc for effiency
% jurka joe and barry for Reliabilityb
% thomas weick teodor for Functionality
% ishii, scrappy, darcy for maintainability

%\todo{Now look at the authors, what are they into?}

% tgl == momjian
% > nauthors
%                     momjian     thomas      davec      neilc      ishii darcy
% None            0.003       0.150       0.048       0.020       0.081       0.15
% Portability     0.211       0.083       0.177       0.173       0.122       0.20
% Efficiency      0.114       0.100       0.080       0.152       0.132       0.00
% Reliability     0.149       0.100       0.137       0.159       0.102       0.15
% Functionality   0.220       0.300       0.266       0.145       0.214       0.20
% Maintainability 0.132       0.150       0.145       0.104       0.224       0.20
% Usability       0.167       0.116       0.145       0.243       0.122       0.10
%                      barry      wieck      inoue      pgsql teodor jurka
% None            0.03488372 0.04444444 0.04166667 0.06521739 0.0250  0.04
% Portability     0.13372093 0.17777778 0.09722222 0.19565217 0.2250  0.18
% Efficiency      0.13372093 0.08888889 0.26388889 0.15217391 0.1250  0.12
% Reliability     0.19767442 0.11111111 0.12500000 0.06521739 0.0875  0.24
% Functionality   0.20930233 0.28888889 0.23611111 0.19565217 0.2375  0.16
% Maintainability 0.13372093 0.06666667 0.13888889 0.17391304 0.1250  0.14
% Usability       0.15697674 0.22222222 0.09722222 0.15217391 0.1750  0.12
%                         tgl    scrappy     petere     dennis     meskes
% None            0.006704981 0.08333333 0.02296451 0.06451613 0.02298851
% Portability     0.207854406 0.25000000 0.20250522 0.22580645 0.23754789
% Efficiency      0.119731801 0.08333333 0.09812109 0.12903226 0.08045977
% Reliability     0.150383142 0.16666667 0.12734864 0.09677419 0.14559387
% Functionality   0.217432950 0.12500000 0.20668058 0.12903226 0.22222222
% Maintainability 0.129310345 0.20833333 0.14196242 0.00000000 0.10727969
% Usability       0.168582375 0.08333333 0.20041754 0.35483871 0.18390805
%                        joe
% None            0.06122449
% Portability     0.22448980
% Efficiency      0.08163265
% Reliability     0.22448980
% Functionality   0.12244898
% Maintainability 0.10204082
% Usability       0.18367347

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\textwidth]{figures/author-distance-from-aggregate}
  \caption{PostgreSQL Author Similarity to the Global Distribution of NFR topics}
\label{fig:authorsim}
\end{figure}

@@

\emph{External validity} -- %was addressed by our use of two case studies, but there are still issues. 
Our data originated from OSS database projects and thus might not be applicable to commercially developed software or other domains. 
Furthermore, our analysis techniques rely on a project's use of meaningful commit messages, although we feel this is the most frequently occurring form
of developer comments. 
%The OSS projects investigated were database systems, thus 
%our results might not generalize to other domains. %We only studied database systems.
